Representation Project

### 2024-11-20 nb/overview.qmd
update with some results

---
title: "Effect of Reference Class on Planning to Conserve"
author:
- name: Thomas E. Gorman
  url: https://tegorman13.github.io/
  affiliations: 
  - name:  Communication and Cognition Lab, Purdue University, USA
    affiliation-url: https://web.ics.purdue.edu/~treimer/
#lightbox: true
toc: false
code-fold: true
code-tools: true
execute:
  #echo: false
  warning: false
format:
  html: 
    grid:
      sidebar-width: 220px
      body-width: 1200px
      margin-width: 170px
      gutter-width: 1.0rem
  hugo-md:
    include: true
    html-math-method: mathjax
    output-file: overview_hugo.md
  gfm:
    echo: true
    output-file: overview_gfm.md
---

@@

pacman::p_load(dplyr,purrr,tidyr,here,tibble,brms,rstan,bayestestR,emmeans,tidybayes,modelsummary,
               ggplot2,gt,knitr,kableExtra,ggh4x,lme4,flextable,pander)
options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)
walk(c("fun_plot"), ~ source(here::here(paste0("scripts/", .x, ".R"))))
theme_set(theme_nice())

s1 <- readRDS(here::here("data/s1_processed.rds")) |> filter(!(id %in% readRDS(here::here("data/s1_discrep_ids.rds"))))
s2_long <- readRDS(here::here("data/s2_processed.rds")) |> filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds"))))

@@

@@

### Study 2 - Subject counts 
```{r}

@@

@@

@@

# binomial regression on trial level data

s1_br <- brm(
    matched_goal ~ refClass + (1|id),
    data = s1_agg,
    family = bernoulli(),
    cores=4,
    iter=2000,
    control = list(adapt_delta = 0.99),
    file=paste0(here::here("data/model_cache",'s1_br.rds')),
    )

@@

@@

```{r}
s1_glm <- glm(matched_goal ~ refClass, data = s1_agg, family = binomial())
summary(s1_glm)

@@

# binomial regression on trial level data

@@

s2_glm <- glm(matched_goal ~ refClass + rounded + pct_goal, data = s2_agg1, family = binomial())
summary(s2_glm)

>>>>

---
title: "Results Overview"
execute:
  #echo: false
  warning: false
format:
  html: 
    grid:
      sidebar-width: 220px
      body-width: 1200px
      margin-width: 170px
      gutter-width: 1.0rem
  hugo-md:
    include: true
    html-math-method: mathjax
    output-file: overview_hugo.md
  gfm:
    echo: true
    output-file: overview_gfm.md
---

@@

pacman::p_load(dplyr,purrr,tidyr,stringr,here,tibble,ggplot2,gt,knitr,kableExtra,ggh4x,patchwork,lme4,flextable,pander)

options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)
walk(c("fun_plot"), ~ source(here::here(paste0("scripts/", .x, ".R"))))
theme_set(theme_nice())

s1 <- readRDS(here::here("data/s1_processed.rds")) |> 
  filter(!(id %in% readRDS(here::here("data/s1_discrep_ids.rds")))) |> 
  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))

s2_long <- readRDS(here::here("data/s2_processed.rds")) |> 
  filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds")))) |> 
  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))

@@

### Study 1 Subject Counts

```{r}
#| label: tbl-s1-counts
# total subjects
n_s1 <- s1 |> ungroup() |> summarise(n=n_distinct(id))
n_s1_removed <-  readRDS(here::here("data/s1_discrep_ids.rds")) |> length()

print(paste("Study 1: Total subjects:", n_s1$n, "Removed subjects:", n_s1_removed))

@@

### Study 2 - Subject counts

n_s2 <- s2_long |> ungroup() |> summarise(n=n_distinct(id))
n_s2_removed <-  readRDS(here::here("data/s2_discrep_ids.rds")) |> length()

print(paste("Study 2: Total subjects:", n_s2$n, "Removed subjects:", n_s2_removed))

@@

### Study 2 Performance

@@

## Group comparison via proportions

```{r}
#| label: tbl-s1-agg

s1_agg <- s1 |> 
    filter(appliance !="Total kWh") |> 
    group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
    summarise(total_kWh = sum(value),orig_kWh=sum(family), 
                pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), 
                n_change = sum(value!=family),
                state_p_dif=mean(state_p_dif),
                state_f_dif=mean(state_f_dif),
                n_less_avg = sum(less_avg),
                duration=first(Duration__in_seconds_)) |> 
    mutate(matched_goal = (pct_change == pct_goal), 
                error = pct_change - pct_goal,
                abs_error = abs(error),
                close_match = abs_error <= 0.03) |>
    ungroup() |> # Add ungroup here
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.15 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        ) |> relocate(accuracy_level, .after= "pct_change")

s1_agg4 <- s1_agg |> group_by(id,refClass,calc) |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(mg=sum(matched_goal),n=n(), pct=mg/n,mean_pct_change=mean(pct_change),mean_abs_error=mean(abs_error),n_accuracy=first(n_accuracy)) |> 
    mutate(accuracy_level = factor(
            case_when(
                mean_abs_error < 0.02 ~ "Exact match",
                mean_abs_error <= 0.02 ~ "0.01-2% error",
                mean_abs_error <= 0.15 ~ "2.01-15% error",
                TRUE ~ "Over 15% error"  # Capture all remaining cases
            ), 
            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
            ordered = TRUE
        ))

# s1_agg |> arrange(id,accuracy_level) |> 
#   select(id,refClass,state,block,plan,total_kWh,orig_kWh,pct_change,accuracy_level,n_change,matched_goal,error,abs_error) |> print(n=20)

observed_props <- s1_agg |>
  group_by(refClass, accuracy_level) |>
  summarise(n = n()) |>
  group_by(refClass) |>
  mutate(prop = n/sum(n)) |>
  pivot_wider(
    names_from = accuracy_level,
    values_from = c(n, prop)
  )

prop_acc_s1 <- s1_agg %>%
    group_by(refClass, accuracy_level) %>%
    summarise(count = n()) %>%
    group_by(refClass) %>%
    mutate(Probability = count / sum(count)) %>%
    ungroup()

# compute percentage of subjects per accuracy level per group
observed_props_s1 <- s1_agg |>
  group_by(refClass, accuracy_level) |>
  summarise(n = n()) |>
  group_by(refClass) |>
  mutate(prop = n/sum(n)) |>
  mutate(n_prop=paste0(n," (",round(prop*100,1),"%)" ), pct_grp=paste0(round(prop*100,1), "%")) |> ungroup()

observed_props_s1 |> 
  mutate(n_total=sum(n)/4) |> 
  group_by(accuracy_level) |>
  mutate(ns=sum(n)/4) |> 
  mutate(Total = paste0(round(ns/n_total*100,1), "%")) |>
  select('Reference Class'=refClass, 'Accuracy Level'=accuracy_level, '% in Group'=pct_grp, "Combined Groups %" =Total) |>
  pivot_wider(
    names_from = 'Reference Class',
    values_from = c('% in Group')
  ) |> relocate("Combined Groups %" , .after=last_col()) |> 
  pander::pandoc.table(caption="Study 1: Proportion of participants per accuracy level by condition")

# observed_props_s1 |> 
#   select('Reference'=refClass, 'Accuracy Level'=accuracy_level, 'n (% group)'=n_prop) |>
#   pivot_wider(
#     names_from = 'Accuracy Level',
#     values_from = c('n (% group)')
#   )

# % of accuracy_level
prop_acc_s1 <- s1_agg %>%
    group_by(refClass, accuracy_level) %>%
    summarise(count = n()) %>%
    group_by(accuracy_level) |>
    mutate(Probability = count / sum(count)) %>%
    ungroup()

# % of entire sample 
prop_acc_s1 <- s1_agg %>%
    group_by(refClass, accuracy_level) %>%
    summarise(count = n()) %>%
    ungroup() |> 
    mutate(Probability = count / sum(count)) %>%
    ungroup()

::: {.content-visible when-format="html"}

```{r}
#| label: tbl-s1-interactive

s1_agg4 |> 
    DT::datatable(extensions = 'Buttons',
    options = list(
       dom = 'Blfrtip',
       buttons = list('copy', 'csv', 'excel', 'pdf', 'print'),
      pageLength = 25,
      autoWidth = TRUE
    ),
    filter = 'top', caption="Exp 1. Participant global averages")

s1_agg |> 
  select(id,refClass,calc,state,plan,pct_change,accuracy_level,n_change,matched_goal,abs_error,total_kWh,orig_kWh) |> 
    group_by(id,refClass,calc) |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    mutate(mean_abs_error=mean(abs_error), .after=abs_error) |> 
    mutate(across(where(is.numeric), \(x) round(x, 3))) %>%
    DT::datatable(extensions = 'Buttons',
    options = list(
       dom = 'Blfrtip',
       buttons = list('copy', 'csv', 'excel', 'pdf', 'print'),
      pageLength = 25,
      autoWidth = TRUE
    ),
    filter = 'top', caption="Exp 1. Participant trial level performance")

s1_agg |> filter(id==124) |> 
  select(id,refClass,state,block,plan,total_kWh,orig_kWh,pct_change,accuracy_level,n_change,matched_goal,abs_error) |> 
  group_by(id) |> mutate(mean_abs_error=mean(abs_error), .after=abs_error) |> 
  pander::pandoc.table(caption="Study 1: Example of participant performance (sbj 124")

:::

```{r}
#| label: tbl-s2-agg
#| fig-width: 12

s2_agg1 <- s2_long |> 
  filter(appliance != "TOTAL") |> 
  group_by(id,refClass,calc, state,pct,pct_goal,plan,rounded) |> 
  summarise(
    total_kWh = sum(value),
    orig_kWh = sum(family),
    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
    state_dif = mean(state_dif),
    .groups = "drop"
  ) |>
  mutate(matched_goal = (pct_change == pct),
                error = pct_change - pct,
                abs_error = abs(error)) |> 
      ungroup() |> # Add ungroup here
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.15 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        )

s2_agg4 <- s2_agg1 |> group_by(id,refClass,calc) |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(mg=sum(matched_goal),n=n(), pct=mg/n,mean_pct_change=mean(pct_change),mean_abs_error=mean(abs_error),n_accuracy=first(n_accuracy)) |> 
    mutate(accuracy_level = factor(
            case_when(
                mean_abs_error < 0.02 ~ "Exact match",
                mean_abs_error <= 0.02 ~ "0.01-2% error",
                mean_abs_error <= 0.15 ~ "2.01-15% error",
                TRUE ~ "Over 15% error"  # Capture all remaining cases
            ), 
            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
            ordered = TRUE
        ))

# s1_agg |> arrange(id,accuracy_level) |> 
#   select(id,refClass,state,block,plan,total_kWh,orig_kWh,pct_change,accuracy_level,n_change,matched_goal,error,abs_error) |> print(n=20)

# compute percentage of subjects per accuracy level per group
observed_props <- s2_agg1 |>
  group_by(refClass, accuracy_level) |>
  summarise(n = n()) |>
  group_by(refClass) |>
  mutate(prop = n/sum(n)) |>
  mutate(n_prop=paste0(n," (",round(prop*100,1),"%)" ))

# confirm that the proportions are with respect to the total number of subjects in each refClass
# 74/.268 = 276; 69*4=276;  38/.138=275;  .435 + .0797 + .21 + .275 = 1

# compute percentage of subjects per accuracy level per group
observed_props_s2 <- s2_agg1 |>
  group_by(refClass, accuracy_level) |>
  summarise(n = n()) |>
  group_by(refClass) |>
  mutate(prop = n/sum(n)) |>
  mutate(n_prop=paste0(n," (",round(prop*100,1),"%)" ), pct_grp=paste0(round(prop*100,1), "%")) |> ungroup()

observed_props_s2 |> 
  mutate(n_total=sum(n)/4) |> 
  group_by(accuracy_level) |>
  mutate(ns=sum(n)/4) |> 
  mutate(Total = paste0(round(ns/n_total*100,1), "%")) |>
  select('Reference Class'=refClass, 'Accuracy Level'=accuracy_level, '% in Group'=pct_grp, "Combined Groups %" =Total) |>
  pivot_wider(
    names_from = 'Reference Class',
    values_from = c('% in Group')
  ) |> relocate("Combined Groups %" , .after=last_col()) |> 
  pander::pandoc.table(caption="Study 1: Proportion of participants per accuracy level by condition")

observed_props_s2 |>  select('Reference'=refClass, 'Accuracy Level'=accuracy_level, 'n (% group)'=n_prop) |>
  pivot_wider(
    names_from = 'Accuracy Level',
    values_from = c('n (% group)')
  )

# % of accuracy_level
prop_acc_s2 <- s2_agg1 %>%
    group_by(refClass, accuracy_level) %>%
    summarise(count = n()) %>%
    group_by(accuracy_level) |>
    mutate(Probability = count / sum(count)) %>%
    ungroup()

# % of entire sample 
prop_combo_s2 <- s2_agg1 %>%
    group_by(refClass, accuracy_level) %>%
    summarise(count = n()) %>%
    group_by(refClass) %>%
    mutate(Probability = count / sum(count)) %>%
    ungroup()

## Study 2 proportion in accuracy plots

p1 <- ggplot(prop_combo_s2, aes(x = accuracy_level, y = Probability, fill = refClass)) +
    geom_bar(stat = "identity", position = position_dodge()) +
  scale_y_continuous(labels = scales::percent) +
    labs(title = "Within-Group %'s per Accuracy bin",
        x = "Accuracy Level",
        y = "Percentage of Participants",
        fill = "Reference Class") +
    theme_minimal()

 p2 <- ggplot(data = s2_agg1, aes(x = accuracy_level, fill = refClass)) +
  geom_bar(position = "dodge", aes(y = (..count..)/sum(..count..))) +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Combined Participant %'s per bin",
       x = "Accuracy Level",
       y = "Percentage of Participants",
       fill = "Reference Class") +
  theme_minimal()

 p3 <- ggplot(prop_acc_s2, aes(x = accuracy_level, y = Probability, fill = refClass)) +
    geom_bar(stat = "identity", position = position_dodge()) +
  scale_y_continuous(labels = scales::percent) +
    labs(title = "Aggregate by Accuracy Bin",
        x = "Accuracy Level",
        y = "Percentage of Participants",
        fill = "Reference Class") +
    theme_minimal()

 # combined plots, and unify axes and legends
{ p1 + p2 +p3 + plot_layout(guides = 'collect') & theme(legend.position = 'right') } + plot_annotation(title = "Study 2: Accuracy Level Proportions")

@@

```{r}
#| eval: false
#| include: false

pacman::p_load(brms,rstan,bayestestR,emmeans,tidybayes,modelsummary)

# binomial regression on trial level data
s1_br <- brm(
    matched_goal ~ refClass + (1|id),
    data = s1_agg,
    family = bernoulli(),
    cores=4,
    iter=2000,
    control = list(adapt_delta = 0.99),
    file=paste0(here::here("data/model_cache",'s1_br.rds')),
    )

@@

```{r}
#| eval: false
#| include: false

@@

```{r}
#| eval: false
#| include: false

s1_glm <- glm(matched_goal ~ refClass, data = s1_agg, family = binomial())
summary(s1_glm)

@@

```{r}
#| eval: false
#| include: false
# binomial regression on trial level data

@@

```{r}
#| eval: false
#| include: false
#| 
s2_glm <- glm(matched_goal ~ refClass + rounded + pct_goal, data = s2_agg1, family = binomial())
summary(s2_glm)

### 2024-11-20 paper.qmd
update with some results

@@

<iframe src="/manuscript/output/manuscript.html" frameborder="0" style="overflow:hidden;height:88%;width:100%;border:2px solid #ccc;" height="88%" width="100%" name="manuscript"></iframe>
```

>>>>

  <a class="btn btn-warning btn-sm mb-1" target="_blank" href="/Manuscript/output/manuscript.docx">
  <iconify-icon inline="" icon="fa6-solid:file-arrow-down"></iconify-icon> &ensp;docx <small>(download)</small>
  </a>

@@

<iframe src="/manuscript/output/manuscript.html" frameborder="0" style="overflow:hidden;height:75%;width:95%;border:2px solid #ccc;" height="75%" width="95%" name="manuscript"></iframe>
```
